---
title: "Descriptive Statistics EU vs EEU"
author: "Anton Yeshchenko, Kerwin Jorbina"
date: "24 April 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Descriptive statistics

### We have a set that weights 500mb, and contains 1421 variables for every country. So we expect to spend quite some time on 'feature engineering', what is the best indicators of country's growth and betterness?


First, we start from looking into GDP, as the main indicator of economy growth. 

```{r, echo=FALSE,include=FALSE, cache=FALSE}
setwd("C:/Users/Anton Yeshchenko/Desktop/Data mining/project")
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape)
library(scales)

dev <- read.csv("Indicators.csv")

eu2004 <- c("Hungary","Slovenia","Slovak Republic","Poland","Lithuania","Latvia","Estonia","Czech Republic")

#these for further analysis as large enough
eu2004_1 <- c("Hungary","Slovak Republic","Poland","Czech Republic")

eu2007 <- c("Bulgaria","Romania")

ukraine <- c("Ukraine")

eeu2010 <- c("Russian Federation","Belarus","Kazakhstan","Armenia","Kyrgyz Republic")

```

## EU countries (joined after 2004)

```{r gdp, echo=FALSE}

plotCountriesGDP <- function(c) {
  compCountry <- c#countries to compare
  countryInd <- dev %>% filter(dev$CountryName %in% compCountry) %>% select(5,2,3,4,6)
  countryInd1 <- countryInd %>% select(1,2,4,5) %>% spread(key = IndicatorCode, value = Value)
  titleformat <- theme(plot.title = element_text(color = "blue", size = 16, face = "bold"))
  
  gdp2004 <- countryInd1 %>% select(Year, CountryCode, one_of(c("NY.GDP.PCAP.PP.CD"))) %>%
    filter(Year >=1995)
  gdp2004
  colnames(gdp2004)[3] <- c("GDP PPP")
  gdp20041 <- melt(gdp2004, id = c("Year", "CountryCode"))
  plot2 <- ggplot(gdp20041, aes(Year, value, col = CountryCode)) +
    geom_line(size = 1.3) + 
    facet_grid(.~variable) + 
   # scale_y_continuous(breaks=seq(0, 20, 2))+
    ylab("GDP numbers") +
    ggtitle("GDP per capita, PPP") +
    theme(axis.title.x = element_blank(),legend.justification = c(1,1), legend.position= c(0.2,1.1),legend.title=element_blank()) +
    titleformat 
  return(plot2)
}

plotCountriesGDP(eu2004)

```

## EU countries (joined after 2007) + Ukraine

```{r gdp07, echo=FALSE}
plotCountriesGDP(c(eu2007,ukraine))
```

## EEU countries


```{r gdp-eeu, echo=FALSE}
plotCountriesGDP(eeu2010)
```

### Now all plots on the same scale, but colored by EU2004-EU2007-EEU

```{r gdp-All, echo=FALSE}
 plotCountriesGDP <- function(eu4,eu7,eeu,ukr) {
  compCountry <- c(eeu,eu4,eu7,ukr)#countries to compare
  countryInd <- dev %>% filter(dev$CountryName %in% compCountry) %>% select(5,2,3,4,6)
  countryInd1 <- countryInd %>% select(1,2,4,5) %>% spread(key = IndicatorCode, value = Value)
  titleformat <- theme(plot.title = element_text(color = "blue", size = 16, face = "bold"))

  gdp2004 <- countryInd1 %>% select(Year, CountryCode, one_of(c("NY.GDP.PCAP.PP.CD"))) %>%
    filter(Year >=1995)
  gdp2004
  colnames(gdp2004)[3] <- c("GDP PPP")
  gdp20041 <- melt(gdp2004, id = c("Year", "CountryCode"))

  plot2 <- ggplot(gdp20041, aes(Year, value, col = CountryCode)) +
    geom_line(size = 1.3) + 
    facet_grid(.~variable) + 
   # scale_y_continuous(breaks=seq(0, 20, 2))+
    ylab("GDP numbers") +
    ggtitle("GDP per capita (EU-EEU-Ukr compared), PPP") +
    theme(axis.title.x = element_blank(),legend.justification = c(1,1), legend.position= c(0.2,1.1),legend.title=element_blank()) +
    titleformat + scale_colour_manual(values = c('purple','yellow','purple','yellow','yellow','purple','purple','yellow', 'yellow','purple','yellow', 'springgreen3'))
  return(plot2)
}

plotCountriesGDP(eu2004_1,eu2007, eeu2010, ukraine)

sort(c(eu2004_1,eu2007, eeu2010, ukraine))

```


### Now some unemployment)


```{r unemployment, echo=FALSE}

compCountry <- c(eu2004_1, eeu2010, ukraine, eu2007)#countries to compare
countryInd <- dev %>% filter(dev$CountryName %in% compCountry) %>% select(5,2,3,4,6)
countryInd1 <- countryInd %>% select(1,2,4,5) %>% spread(key = IndicatorCode, value = Value)
titleformat <- theme(plot.title = element_text(color = "blue", size = 16, face = "bold"))



unemployment <- countryInd1 %>% select(Year, CountryCode, one_of(c("SL.UEM.TOTL.ZS", "SL.UEM.TOTL.FE.ZS", "SL.UEM.TOTL.MA.ZS"))) %>%
  filter(Year >=1995)
colnames(unemployment)[3] <- c("Total", "Female", "Male")
unemployment1 <- melt(unemployment, id = c("Year", "CountryCode"))
plot2 <- ggplot(unemployment1, aes(Year, value, col = CountryCode)) +
  geom_line(size = 1.5) + 
  facet_grid(.~variable) + 
  scale_y_continuous(breaks=seq(0, 40, 1))+
  ylab("Unemployment (% of total") +
  ggtitle("UNEMPLOYMENT LEVEL") +
  theme(axis.title.x = element_blank(),legend.justification = c(1,1), legend.position= c(1,1),legend.title=element_blank()) +
  titleformat      
plot2
```

